---
title: 'EDS 223: assignment 1'
author: '{student name and collaborators}'
date: "2022-09-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## introduction
The following exercises are modified from [Chapter 9 of Geocomputation with R by Rovin Lovelace](https://geocompr.robinlovelace.net/adv-map.html). Each question lists the total number of points. The breakdown of points can be found at the end of each instruction in parentheses. A general grading rubric can be found on the [course website](https://ryoliver.github.io/EDS_223_spatial_analysis/assignments.html).

**Please update "author" to list your first and last names and any collaborators (e.g. Ruth Oliver, Friend1, Friend2)**

**Due by midnight Saturday 2022-10-08**

## prerequisites

```{r load, include=TRUE, message=FALSE, warning=FALSE}
library(sf)
library(raster)
library(dplyr)
library(spData)
library(spDataLarge)
library(tmap)
```

These exercises rely on a new data object based on the **world** and **worldbank_df** datasets from the **spData* package.

```{r include=TRUE}
africa = world |> 
  filter(continent == "Africa", !is.na(iso_a2)) |> 
  left_join(worldbank_df, by = "iso_a2") |> 
  dplyr::select(name, subregion, gdpPercap, HDI, pop_growth) |> 
  st_transform("+proj=aea +lat_1=20 +lat_2=-23 +lat_0=0 +lon_0=25")
```

We will also use the **zion** and **nlcd** datasets from the **spDataLarge** package.

```{r include=TRUE}
zion = st_read((system.file("vector/zion.gpkg", package = "spDataLarge")))
data(nlcd, package = "spDataLarge")
force(nlcd)
```

## question 1
##### 10 points

Create a map showing the geographic distribution of the Human Development Index (HDI) across Africa.  
-   use both base graphics (hint: use **plot()**) and **tmap**) (4)   
-   name two advantages of each based on the experience (3) 
-   name three other mapping packages and an advantage of each (3)


```{r include=TRUE}
```

## question 2
##### 20 points

Extend the **tmap** created for question 1 so the legend has three bins: "high" (HDI above 0.7), "medium" (HDI between 0.55 and 0.7), and "low" (HDI below 0.55). (5)  
-   change the legend title (5)  
-   change the class labels (5)  
-   change the color palette (5)  

```{r include=TRUE}
```

## question 3
##### 20 points

Represent Africa's subregions on the map. (5)    
-   change the color palette (5)  
-   change the legend title (5)  
-   combine this map with the map from question 2 into a single plot (5)   

```{r include=TRUE}
```

## question 4
##### 30 points

Create a land cover map of Zion National Park (5)    
-   change the default colors to match your perception of land cover categories (5)    
-   move the map legend outside of the map to improve readability (5)    
-   add a scale bar and north arrow and change the position of both to improve the maps aesthetics (5)  
-   add the park boundaries on top of the land cover map (5)    
-   add an inset of Zion's location in the context of the state of Utah (5)    
    -   hint: an object representing Utah can be subset from the **us_states** dataset)  

```{r include=TRUE}
```

## question 5
##### 20 points

Create an interactive map of the world (10)    
-   include a legend (5)  
-   change the color palette (5)  
    -   bonus: use **leaflet** insted of **tmap** (2)   

```{r include=TRUE}
```

## bonus question
##### 5 points
Create THE WORST map! AKA a map that breaks all of the rules of legibility/clarity, yet somehow still passes for a map. We'll vote on the best worst map (worst worst map?) in class.

```{r include=TRUE}
```