---
title: 'EDS 223: assignment 1'
author: "{update with your name and study partners}"
output: html_document
date: "2023-10-03"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview
In this assignment we'll explore making maps in R! To do so, we will be visualizing environmental inequities within the United States.

We will be working with data from the United States Environmental Protection Agency's EJScreen: Environmental Justice Screening and Mapping Tool.

According to the [US EPA website](https://www.epa.gov/ejscreen/purposes-and-uses-ejscreen):

>This screening tool and data may be of interest to community residents or other stakeholders as they search for environmental or demographic information. It can also support a wide range of research and policy goals. The public has used EJScreen in many different locations and in many different ways.
>
>EPA is sharing EJScreen with the public:  
> - to be more transparent about how we consider environmental justice in our work,  
> - to assist our stakeholders in making informed decisions about pursuing environmental justice and,   
> - to create a common starting point between the agency and the public when looking at issues related to environmental justice. 

EJScreen provides on environmental and demographic information for the US at the Census [tract](https://en.wikipedia.org/wiki/Census_tract) and [block group](https://en.wikipedia.org/wiki/Census_block_group) levels. You will be working with block group data that has been downloaded from the [EPA site](https://www.epa.gov/ejscreen/download-ejscreen-data). To understand the associated data columns, you will need to explore the Technical Documentation and column description spreadsheet available in the data folder. I also encourage you to explore the [limitations and caveats](https://www.epa.gov/ejscreen/limitations-and-caveats-using-ejscreen) of the data.


## Assignment
Your task is to explore a question or problem of your choosing. You should focus on region(s), communities, or environmental hazards that interest you.  

Please write a short paragraph describing the environmental question you would like to explore through visualization. Create 3 distinct maps of your choice that support this exploration. Because breaking the rules can be a fun way to learn them, please make the ugliest map you can imagine that breaks all the rules!
  
Please incorporate the following elements into at least **one map**:

-   categorical color display
-   continuous color display
-   custom binned color legend
-   graticules
-   scale bar
-   compass rose
-   interactive map
-   inset map
-   updated color palette
-   display two datasets in one map

**All maps** should include the following:

-   legend with legible (updated) title
-   legible colorscale
-   indication of scale and orientation 
  
### Grading
Assignment 1 is worth 15% of your final grade. Full credit will be given to responses that address all requirements listed above. Partial credit will be given based on the proportion of requirements. The short paragraph is worth 20% of the assignment grade. Maps are worth 80% of the assignment grade.


Below I provide example code to help you manipulate the EJScreen data. You should update these to the regions of your choice.
```{r include=TRUE, eval=FALSE}
library(tidyverse)
library(sf)

# read in geodatabase of EJScreen data at the Census Block Group level
ejscreen <- st_read("~/data/EJSCREEN_2023_BG_StatePct_with_AS_CNMI_GU_VI.gdb/") 

# filter to a state you are interested in
california <- ejscreen %>%
  filter(ST_ABBREV == "CA") 

# filter to a county you are interested in
santa_barbara <- ejscreen %>%
  filter(CNTY_NAME %in% c("Santa Barbara County"))

# find the average values for all variables within counties
california_counties <- aggregate(california, by = list(california$CNTY_NAME), FUN = mean)

# find the centroid within each Census Block Group
# this is helpful so you can plot Census Block Group data as points
california_blockgroup_centroids <- st_centroid(california)

```
